### 2025-05-14 analysis/paper/loan word maps.qmd
network analysis

---
title: "Make interactive loan word maps"
author:
  - Henning Schreiber:
      orcid: 0000-0001-9576-8036
      institute: uhh

  - Nikolas Gestrich:
      correspondence: "yes"
      email: gestrich@uni-frankfurt.de
      institute: fi
      orcid: 0000-0001-6376-1972

institute:
  - uhh:
      name: Asien-Afrika-Institut, Universität Hamburg
      address: Mittelweg 177, 20148 Hamburg
  - fi: 
      name: Frobenius Institute
      address: Norbert-Wollheim-Platz 1, 60323 Frankfurt am Main, Gemany
title-block-published: "Last updated"  
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  West Africa; trade; historical linguistics
highlights: |
  Maps of loanwords to illustrate historical trade pattern in West Africa
prefer-html: true
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

<!-- With the following code you can access and display values from the yml header above. -->

Keywords: `r rmarkdown::metadata$keywords`

Highlights: `r rmarkdown::metadata$highlights`

<!-- The actual document text starts here: -->

## Retrieve and clean data
In order to have location data, we read in the African languages, their glottoglog code, and their location from the glottolog database (@hammarstromGlottolog2024).

```{r}
#| label: get-data
#| eval: true
# Read data
langloc <- readr::read_csv(here::here('analysis/data/raw_data/glottologAfricanLanguages.csv')) |>

  #retain only relevant variables
  dplyr::select(glott_code, latitude, longitude, LanName) |>

# drop languages with unknown location
  tidyr::drop_na(longitude) |>

  # turn latitude and longitude columns into geometry
  sf::st_as_sf(coords = c("longitude","latitude"))

# set crs
sf::st_crs(langloc) <- 4326
```

Next, we read in the data and join the entries to the glottolog entries. Modify the list of .csv files to suit your interests. If you want to upload new files, place these in the directory /analysis/data/raw_data.Make sure any new data has a glottolog language code in a variable named "glott_code"

```{r read data}
# read classified loan word data
# Create a vector of file paths
objs <- here::here("analysis", "data", "raw_data", c("gold.csv", 
                                                     "fonio.csv",
                                                     "millet.csv",
                                                     "salt.csv",
                                                     "iron.csv"))

# read the csv files into a list of data frames
data_list <- purrr::map(objs, readr::read_csv)

# Add the location data to the word lists

# Function to join location data to each data frame
join_geometry <- function(df) {
  df |> 
    dplyr::left_join(langloc, by = "glott_code") |> 
    sf::st_as_sf()  # Convert the result to an sf object if needed
}

# Apply the join function to each data frame in the list
data_list <- purrr::map(data_list, join_geometry)

# Assign names
names(data_list) <- c("gold", "fonio", "millet", "salt", "iron")
```

The data may contain certain duplicate entries. These need to be removed to avoid overly full plots. This is done by running the code below.

```{r clean duplicates}
# Function to remove duplicates
remove_duplicates <- function(df_list) {
  df_list <- lapply(df_list, function(df) {
    df <- df |> 
      dplyr::distinct(glott_code, style, .keep_all = TRUE)
    return(df)
  })
  return(df_list)
}

# Apply function
data_list <- remove_duplicates(data_list)
```

Classifications go into more detail than we need for plotting. We will only retain a rough classification here:

```{r}
data_list <- data_list |> 
  purrr::map(~ dplyr::mutate(.x, style = stringr::str_extract(style, "^[A-Za-z]")))
```

## Make interactive maps

Function to make interactive leaflet maps

```{r map function}
create_leaflet_map <- function(data_sf, form_column = "form", style_column = "style", background = "colour") {

  # Ensure the input data is an sf object and contains the required columns
  if (!inherits(data_sf, "sf")) {
    stop("The input data must be an sf object.")
  }
  if (!style_column %in% colnames(data_sf)) {
    stop(paste("The column", style_column, "is not found in the data."))
  }
  if (!"glott_code" %in% colnames(data_sf)) {
    stop("The column 'glott_code' is not found in the data.")
  }
  if (!"LanName" %in% colnames(data_sf)) {
    stop("The column 'LanName' is not found in the data.")
  }

  # Filter out rows with NA values in the "style" column
  data_sf <- data_sf |>
    dplyr::filter(!is.na(!!rlang::sym(style_column)))

  # Set the tile provider based on the 'background' argument
  tile_provider <- switch(
    background,
    "colour" = leaflet::providers$Esri.WorldPhysical,  # Default color tile
    "bw" = leaflet::providers$Esri.WorldGrayCanvas,    # Black-and-white tile
    stop("Invalid value for 'background'. Choose 'colour' or 'bw'.")
  )

  # Define a color palette for the "style" column (using Dark2 palette from RColorBrewer)
  color_pal <- leaflet::colorFactor(palette = RColorBrewer::brewer.pal(8, "Dark2"), domain = data_sf[[style_column]])

  # Add a new column with the corresponding colors based on the 'style' column
  data_sf <- data_sf |>
    dplyr::mutate(color = color_pal(!!rlang::sym(style_column)))  # Dynamically reference the column

  # Adjust overlapping points by slightly jittering their coordinates
  # Initially apply fixed jitter for the map creation
  jitter_scale <- 0.5000  # Set the base jitter scale factor

  data_sf <- data_sf |>
    dplyr::mutate(
      original_lon = sf::st_coordinates(geometry)[, 1],  # Store original longitude
      original_lat = sf::st_coordinates(geometry)[, 2],  # Store original latitude
      jittered_lon = original_lon + stats::rnorm(dplyr::n(), mean = 0, sd = jitter_scale),
      jittered_lat = original_lat + stats::rnorm(dplyr::n(), mean = 0, sd = jitter_scale)
    )

  # Create the leaflet map
  leaflet_map <- leaflet::leaflet(data_sf) |>
    leaflet::addProviderTiles(tile_provider) |>
    leaflet::addLabelOnlyMarkers(
      lng = ~jittered_lon,  # Use jittered longitude
      lat = ~jittered_lat,  # Use jittered latitude
      label = purrr::map(
        glue::glue("<span style='color:{data_sf$color}; font-weight: bold; font-size: 12px;'>{as.character(data_sf[[form_column]])}</span>"),
        htmltools::HTML
      ),
      labelOptions = leaflet::labelOptions(
        noHide = TRUE,
        direction = 'top',
        textOnly = TRUE
      )
    ) |>
    leaflet::addMarkers(
      lng = ~jittered_lon,
      lat = ~jittered_lat,
      popup = ~paste("<strong>Glottocode:</strong> ", glott_code, "<br><strong>Language Name:</strong> ", LanName),
      options = leaflet::markerOptions(opacity = 0)
    )

  # Add JavaScript to dynamically adjust jitter based on zoom level
  leaflet_map <- leaflet_map |>
    htmlwidgets::onRender(
      "
      function(el, map) {
        // Function to apply jitter based on zoom level
        function applyJitter() {
          var zoomLevel = map.getZoom();
          var jitterScale = 0.5 * Math.pow(20, -zoomLevel); // Increase jitter at lower zoom levels

          // Adjust the jitter of each marker
          map.eachLayer(function(layer) {
            if (layer instanceof L.Marker) {
              var originalLat = layer.options.originalLat;  // Access the original latitude
              var originalLng = layer.options.originalLng;  // Access the original longitude
              var newLat = originalLat + (Math.random() - 0.5) * jitterScale;
              var newLng = originalLng + (Math.random() - 0.5) * jitterScale;
              layer.setLatLng([newLat, newLng]);
            }
          });
        }

        // Apply jitter on map zoom change
        map.on('zoomend', function() {
          applyJitter();
        });

        // Apply jitter initially
        applyJitter();
      }
      ",
      list(
        originalLat = ~original_lat,
        originalLng = ~original_lon
      )
    )

  # Return the map object
  return(leaflet_map)
}
```

Function to save a version of every map in the list of data

```{r save function}
# Function to create and save leaflet maps for all data frames in data_list
save_maps <- function(data_list, form_column = "form", background = "colour") {

  # Use purrr::map2 to iterate over the list of data frames and generate maps
  purrr::map2(data_list, seq_along(data_list), ~ {
    # Create the map using the create_leaflet_map function
    map <- create_leaflet_map(.x, form_column, background)

    # Generate the filename based on the index or data frame name
    filename <- paste0("map_", .y, ".html")

    # Save the map as an HTML file in the 'analysis/figures' folder
    htmlwidgets::saveWidget(map, here::here("analysis/figures", filename))
  })
}
```

```{r}
#| label: fonio
#| fig-cap: "fonio word forms"
# create_leaflet_map(data_list$iron, background = "bw") #uncomment to run
```

Save an html version of all objects

```{r}
#| label: save-maps
#| eval: false
# save_maps(data_list)  #uncomment to run
```

### Add further layers to the map
Get the archaeological data layers
```{r}
# Create a vector of file paths
points <- here::here("analysis", "data", "raw_data", c("gold_points.csv", 
                                                     "fonio_points.csv",
                                                     "millet_points.csv",
                                                     "salt_points.csv",
                                                     "iron_points.csv"))

# read the csv files into a list of data frames
point_list <- purrr::map(points, readr::read_csv2)

# make sf geometry
points_sf <- point_list[[1]] |> sf::st_as_sf(coords = c("lon","lat"))

# set crs
sf::st_crs(points_sf) <- 4326
```

function to add further layers
```{r}
add_sf_to_leaflet_map <- function(leaflet_map, new_data_sf, label_column) {

  # Ensure the new input data is an sf object and contains the required columns
  if (!inherits(new_data_sf, "sf")) {
    stop("The new input data must be an sf object.")
  }
  if (!label_column %in% colnames(new_data_sf)) {
    stop(paste("The column", label_column, "is not found in the new data."))
  }

  # Define a color for the new points
  new_point_color <- "black"

  # Add the new sf object as a point layer to the existing leaflet map
  leaflet_map <- leaflet_map |>
    leaflet::addCircleMarkers(
      data = new_data_sf,
      lng = ~sf::st_coordinates(geometry)[, 1],
      lat = ~sf::st_coordinates(geometry)[, 2],
      color = new_point_color,
      radius = 2,
      stroke = FALSE, fillOpacity = 1,
      label = purrr::map(
        glue::glue("<span style='color:{new_point_color}; font-weight: normal; font-size: 10px;'>{as.character(new_data_sf[[label_column]])}</span>"),
        htmltools::HTML
      ),
      labelOptions = leaflet::labelOptions(
        noHide = TRUE,
        direction = 'top',
        textOnly = TRUE
      )
    )

  # Return the updated map object
  return(leaflet_map)
}
```

```{r}
#run mapping function
#create_leaflet_map(data_list$gold, background = "bw") |> #uncomment to run

#add point layer
#add_sf_to_leaflet_map(points_sf, label_column = "site") #uncomment to run
```

## Networks
Next, we will create networks from this data, in order to be able to analyse and visualise connections.The dataset of languages is the nodelist, the edges are made on the basis of shared forms. First, the edge list is made. Fir this, we need more detailed classifications than for the maps above, so we need to read the data in again without simplifying the classifications.

```{r prepare data for network}
# read classified loan word data

# read the csv files into a list of data frames
net_data_list <- purrr::map(objs, readr::read_csv)

# Assign names
names(net_data_list) <- c("gold", "fonio", "millet", "salt", "iron")

# Function to prepare each dataset for making it into an edge list
prepare_data <- function(data) {
  data |> 
    # Step 1: Split the 'style' column into hierarchical levels
    dplyr::mutate(
      cl1 = stringr::str_extract(style, "^[A-Z]"),                    # First level
      cl2 = stringr::str_extract(style, "^[A-Z]\\.\\d+"),             # Second level
      cl3 = stringr::str_extract(style, "^[A-Z]\\.\\d+\\.\\w+")       # Third level
    ) |> 
    # Step 2: Replace "X" with NA in cl1, cl2, and cl3
    dplyr::mutate(
      cl1 = dplyr::if_else(cl1 == "X", NA_character_, cl1),
      cl2 = dplyr::if_else(cl2 == "X", NA_character_, cl2),
      cl3 = dplyr::if_else(cl3 == "X", NA_character_, cl3)
    )
}

# Apply the function to each dataset in the list
net_data_list <- lapply(net_data_list, prepare_data) 
```

```{r edge list function}
# Function to generate edge list
generate_edgelist <- function(data) {
  # Step 1: Create pairwise comparisons of glott_code values
  edgelist <- expand.grid(glott_code_1 = data$glott_code, 
                          glott_code_2 = data$glott_code,
                          stringsAsFactors = FALSE) |>
    # Remove self-loops
    dplyr::filter(glott_code_1 != glott_code_2) |>
    # Add corresponding cl1, cl2, cl3 for both nodes
    dplyr::left_join(data, by = c("glott_code_1" = "glott_code")) |>
    dplyr::rename(cl1_1 = cl1, cl2_1 = cl2, cl3_1 = cl3) |>
    dplyr::left_join(data, by = c("glott_code_2" = "glott_code")) |>
    dplyr::rename(cl1_2 = cl1, cl2_2 = cl2, cl3_2 = cl3) |>
    # Calculate weights based on equivalence
    dplyr::mutate(
      weight = dplyr::case_when(
        cl1_1 == cl1_2 & cl2_1 == cl2_2 & cl3_1 == cl3_2 ~ 3,  # Match in cl1, cl2, and cl3
        cl1_1 == cl1_2 & cl2_1 == cl2_2 ~ 2,                  # Match in cl1 and cl2
        cl1_1 == cl1_2 ~ 1,                                   # Match in cl1 only
        TRUE ~ 0                                              # No match
      )
    ) |>
    # Filter out pairs with weight = 0 (no equivalence)
    dplyr::filter(weight > 0) |>
    dplyr::select(glott_code_1, glott_code_2, weight)

  # Step 2: Reduce the edgelist to retain only one edge per pair with the highest weight
  reduced_edgelist <- edgelist |>
    # Ensure undirected edges by sorting the glott_code pair
    dplyr::mutate(
      node_1 = pmin(glott_code_1, glott_code_2),
      node_2 = pmax(glott_code_1, glott_code_2)
    ) |>
    # Group by the unique undirected pair
    dplyr::group_by(node_1, node_2) |>
    # Retain only the edge with the maximum weight
    dplyr::summarise(weight = max(weight), .groups = "drop") |>
    # Rename columns back to glott_code_1 and glott_code_2 for clarity
    dplyr::rename(glott_code_1 = node_1, glott_code_2 = node_2)

  return(reduced_edgelist)
}
```

```{r create edge lists}
edgelists <- lapply(net_data_list, generate_edgelist)
```

now, we read this into tidygraph network objects. First we take the language list and convert it into a usable edge list, and then make both into a network.
```{r}
# Convert langloc into a tibble with latitude and longitude fields
node_list <- langloc |>
  dplyr::mutate(
    latitude = sf::st_coordinates(geometry)[, 2],
    longitude = sf::st_coordinates(geometry)[, 1]
  ) |>
  tibble::as_tibble() |> 
  dplyr::select(glott_code, LanName, latitude, longitude) |>

#add some missing values
 tibble::add_row(glott_code = "dogo1299", 
                             LanName = "Dogon", 
                             latitude = 14.43, 
                             longitude = -3.49 )

# Create a list of tidygraph objects using the filtered node list
graphs <- lapply(edgelists, function(edges) {
  # Filter the node list to include only nodes present in the edge list
  filtered_nodes <- node_list |>
    dplyr::filter(glott_code %in% c(edges$glott_code_1, edges$glott_code_2))

  # Create the tidygraph object
  tidygraph::tbl_graph(
    nodes = filtered_nodes,
    edges = edges,
    directed = FALSE
  )
})

```

```{r graph plotting functions}
# Function to plot a tidygraph object with geographic layout
plot_geo_graph <- function(graph, latitude = "latitude", longitude = "longitude", weight = "weight", label = "LanName", title = "title") {
  # Ensure that the graph has the required attributes
  if (!latitude %in% colnames(tidygraph::as_tibble(graph, active = "nodes"))) {
    stop(paste("Latitude column", latitude, "not found in node attributes."))
  }
  if (!longitude %in% colnames(tidygraph::as_tibble(graph, active = "nodes"))) {
    stop(paste("Longitude column", longitude, "not found in node attributes."))
  }
  if (!weight %in% colnames(tidygraph::as_tibble(graph, active = "edges"))) {
    stop(paste("Weight column", weight, "not found in edge attributes."))
  }
  if (!label %in% colnames(tidygraph::as_tibble(graph, active = "nodes"))) {
    stop(paste("Label column", label, "not found in node attributes."))
  }

  # Plot the tidygraph object
  ggraph::ggraph(graph, layout = "manual", x = !!rlang::sym(longitude), y = !!rlang::sym(latitude)) +
    ggraph::geom_edge_link(
      ggplot2::aes(width = !!rlang::sym(weight), alpha = !!rlang::sym(weight)), 
      color = "gray"
    ) +
    ggraph::scale_edge_alpha_continuous(range = c(0.6, 1)) + # Map weight to opacity
    ggraph::scale_edge_width_continuous(range = c(0.6, 1)) + # Map weight to width
    ggraph::geom_node_point(size = 5, color = "blue") +
    ggraph::geom_node_text(
      ggplot2::aes(label = !!rlang::sym(label)), 
      repel = TRUE, 
      size = 3, 
      color = "black"
    ) +
    ggplot2::labs(title = title) +
    ggraph::theme_graph() +
    ggplot2::theme(
      legend.position = "none"
    )
}

# Function to plot a tidygraph object with force-directed layout
plot_fr_graph <- function(graph, latitude = "latitude", longitude = "longitude", weight = "weight", label = "LanName", title = "title") {
  # Ensure that the graph has the required attributes
  if (!weight %in% colnames(tidygraph::as_tibble(graph, active = "edges"))) {
    stop(paste("Weight column", weight, "not found in edge attributes."))
  }
  if (!label %in% colnames(tidygraph::as_tibble(graph, active = "nodes"))) {
    stop(paste("Label column", label, "not found in node attributes."))
  }

  # Plot the tidygraph object
  ggraph::ggraph(graph, layout = "fr") +
    ggraph::geom_edge_link(
      ggplot2::aes(width = !!rlang::sym(weight), alpha = !!rlang::sym(weight)), 
      color = "gray"
    ) +
    ggraph::scale_edge_alpha_continuous(range = c(0.6, 1)) + # Map weight to opacity
    ggraph::scale_edge_width_continuous(range = c(0.6, 1)) + # Map weight to width
    ggraph::geom_node_point(size = 5, color = "blue") +
    ggraph::geom_node_text(
      ggplot2::aes(label = !!rlang::sym(label)), 
      repel = TRUE, 
      size = 3, 
      color = "black"
    ) +
    ggplot2::labs(title = title) +
    ggraph::theme_graph() +
    ggplot2::theme(
      legend.position = "none"
    )
}

# Function to plot a tidygraph object with geographic layout and louvain communities
plot_comm_graph <- function(graph, latitude = "latitude", longitude = "longitude", weight = "weight", label = "LanName", title = "title", subtitle = NULL, community = "louvain") {
  # Ensure that the graph has the required attributes
  if (!latitude %in% colnames(tidygraph::as_tibble(graph, active = "nodes"))) {
    stop(paste("Latitude column", latitude, "not found in node attributes."))
  }
  if (!longitude %in% colnames(tidygraph::as_tibble(graph, active = "nodes"))) {
    stop(paste("Longitude column", longitude, "not found in node attributes."))
  }
  if (!weight %in% colnames(tidygraph::as_tibble(graph, active = "edges"))) {
    stop(paste("Weight column", weight, "not found in edge attributes."))
  }
  if (!label %in% colnames(tidygraph::as_tibble(graph, active = "nodes"))) {
    stop(paste("Label column", label, "not found in node attributes."))
  }
   if (!label %in% colnames(tidygraph::as_tibble(graph, active = "nodes"))) {
    stop(paste("Label column", community, "not found in node attributes."))
  }

  # Plot the tidygraph object
  ggraph::ggraph(graph, layout = "manual", x = !!rlang::sym(longitude), y = !!rlang::sym(latitude)) +
    ggraph::geom_edge_link(
      ggplot2::aes(width = !!rlang::sym(weight), alpha = !!rlang::sym(weight)), 
      color = "gray"
    ) +
    ggraph::scale_edge_alpha_continuous(range = c(0.6, 1)) + # Map weight to opacity
    ggraph::scale_edge_width_continuous(range = c(0.6, 1)) + # Map weight to width
    ggraph::geom_node_point(aes(colour = !!rlang::sym(community)),size = 5) +
    ggraph::geom_node_text(
      ggplot2::aes(label = !!rlang::sym(label)), 
      repel = TRUE, 
      size = 3, 
      color = "black"
    ) +
    ggplot2::labs(title = title, subtitle = subtitle) +
    ggraph::theme_graph() +
    ggplot2::theme(
      legend.position = "none"
    )
}

```

```{r}
plot_geo_graph(graphs$gold, title = "gold")
```

Improve the gold graph:
remove Wolof. It has the same form as the eastern Group, but its ouro form is probably portuguese influenced, whereas the lower Niger is not.
Retain only stronger ties.

```{r}
library(tidygraph)
graphs$gold |> 
  mutate(louvain = group_louvain(weights = weight)) |> 
  mutate(louvain = as.character(louvain)) |> 
  activate(edges) |> 
  filter(weight > 1) |> 
  activate(nodes) |>
  filter(!node_is_isolated()) |>

  plot_comm_graph(community = "louvain", title = "gold", subtitle = "Louvain network communities")

```

```{r}
graphs$salt |> 
  mutate(louvain = group_louvain(weights = weight)) |> 
  mutate(louvain = as.character(louvain)) |> 
  activate(edges) |> 
  filter(weight > 1) |> 
  activate(nodes) |>
  filter(!node_is_isolated()) |>

  plot_comm_graph(community = "louvain", title = "salt", subtitle = "Louvain network communities")
```

Create a graph each with betweenness centrality
```{r betweenness graphs}
library(tidygraph)
library(ggraph)
library(dplyr)

# Iterate over each "type" in the edge list and create a graph
edge_list_summary |>
  dplyr::group_split(type) |>
  purrr::walk(\(edges_for_type) {
    # Get the type name
    type_name <- unique(edges_for_type$type)

    # Create a filtered tidygraph object for this type
    graph_for_type <- tidygraph::tbl_graph(
      nodes = node_list,
      edges = edges_for_type,
      directed = FALSE
    ) |>
      tidygraph::filter(!tidygraph::node_is_isolated()) # Remove isolated nodes

    # Compute betweenness centrality for node sizes
    graph_for_type <- graph_for_type |>
      tidygraph::mutate(betweenness = tidygraph::centrality_betweenness())

    # Plot the graph
    plot <- ggraph::ggraph(graph_for_type, layout = "manual", x = longitude, y = latitude) +
      ggraph::geom_edge_link(aes(width = weight), edge_colour = "grey") +
      ggraph::geom_node_point(aes(size = betweenness), colour = "blue") +
      ggraph::geom_node_text(aes(label = LanName), repel = TRUE, size = 3) +
      ggplot2::theme_void() +
      ggplot2::labs(
        title = paste("Graph for Type:", type_name),
        size = "Betweenness",
        width = "Edge Weight"
      ) +
      ggplot2::scale_size_continuous(range = c(2, 10)) +
      ggraph::scale_edge_width_continuous(range = c(0.5, 3)) +
      ggraph::theme_graph(base_size = 12)

    print(plot)
  })
```
Create a graph each with eigenvector centrality
```{r betweenness graphs}
library(tidygraph)
library(ggraph)
library(dplyr)

# Iterate over each "type" in the edge list and create a graph
edge_list_summary |>
  dplyr::group_split(type) |>
  purrr::walk(\(edges_for_type) {
    # Get the type name
    type_name <- unique(edges_for_type$type)

    # Create a filtered tidygraph object for this type
    graph_for_type <- tidygraph::tbl_graph(
      nodes = node_list,
      edges = edges_for_type,
      directed = FALSE
    ) |>
      tidygraph::filter(!tidygraph::node_is_isolated()) # Remove isolated nodes

    # Compute betweenness centrality for node sizes
    graph_for_type <- graph_for_type |>
      tidygraph::mutate(eigenvector = tidygraph::centrality_eigen())

    # Plot the graph
    plot <- ggraph::ggraph(graph_for_type, layout = "manual", x = longitude, y = latitude) +
      ggraph::geom_edge_link(aes(width = weight), edge_colour = "grey") +
      ggraph::geom_node_point(aes(size = eigenvector), colour = "purple") +
      ggraph::geom_node_text(aes(label = LanName), repel = TRUE, size = 3) +
      ggplot2::theme_void() +
      ggplot2::labs(
        title = paste("Graph for Type:", type_name),
        size = "Eigenvector",
        width = "Edge Weight"
      ) +
      ggplot2::scale_size_continuous(range = c(2, 10)) +
      ggraph::scale_edge_width_continuous(range = c(0.5, 3)) +
      ggraph::theme_graph(base_size = 12)

    print(plot)
  })
```

Save a graph for each of the trade items
```{r}
# Create a list of tidygraph objects, one for each "type"
graph_list <- edge_list_summary |>
  dplyr::group_split(type) |>
  purrr::map(\(edges_for_type) {
    # Get the type name
    type_name <- unique(edges_for_type$type)

    # Create a filtered tidygraph object for this type
    graph_for_type <- tidygraph::tbl_graph(
      nodes = node_list,
      edges = edges_for_type,
      directed = FALSE
    ) |>
      tidygraph::filter(!tidygraph::node_is_isolated()) |> # Remove isolated nodes
      tidygraph::mutate(betweenness = tidygraph::centrality_betweenness()) # Add betweenness centrality

    # Return the graph, named with the type
    graph_for_type
  }) |>
  purrr::set_names(unique(edge_list_summary$type))

# Save the list of graphs for further processing
save(graph_list, file = "graph_list.RData")

```

```{r}

```

```{r}
#knock out WOlof from the gold network
graph |> 
  activate(edges) |> 
  filter(type == "gold") |> 
  activate(nodes) |> 
  filter(LanName != "Wolof") |>
  filter(!node_is_isolated()) |> 
  plot_graph()
```

The graphs need to be sparsified now, as they are relatively useless as they are.
```{r}
library(backbone)

disp <- disparity(graph_list$gold, alpha = 0.5)
plot_graph(disp)

```

```{r}
# Find glott_code values in edge lists that are not in the node_list
find_missing_glott_codes <- function(edgelists, node_list) {
  # Combine all unique glott_code values from the edge lists
  all_glott_codes <- lapply(edgelists, function(edges) {
    unique(c(edges$glott_code_1, edges$glott_code_2))
  }) |>
    unlist() |>  # Flatten into a single vector
    unique()     # Remove duplicates

  # Find glott_codes not present in the node_list
  missing_glott_codes <- setdiff(all_glott_codes, node_list$glott_code)

  return(missing_glott_codes)
}

# Apply the function
missing_glott_codes <- find_missing_glott_codes(edgelists, node_list)

# Display the missing glott_code values
print(missing_glott_codes)

```

<!-- The following line inserts a page break  -->

\newpage

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r}
#| label: colophon
#| cache: false

# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```